---
title: "Practical Machine Learning Assignment"
author: "B. Vink"
date: "26 april 2016"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

## 1. Introduction
Using devices such as *Jawbone Up*, *Nike FuelBand*, and *Fitbit* it is now 
possible to collect a large amount of data about personal activity relatively 
inexpensively. These type of devices are part of the quantified self movement 
– a group of enthusiasts who take measurements about themselves regularly to 
improve their health, to find patterns in their behavior, or because they are 
tech geeks. One thing that people regularly do is quantify **how much** of a 
particular activity they do, but they rarely quantify **how well** they do it. 

This project, will use data from accelerometers on the belt, forearm, arm, and 
dumbell of 6 participants. They were asked to perform barbell lifts correctly 
and incorrectly in 5 different ways.  The five ways are:

- **Class A**:   Exactly according to the specification 
- **Class B**:   Throwing the elbows to the front 
- **Class C**:   Lifting the dumbbell only halfway 
- **Class D**:   Lowering the dumbbell only halfway 
- **Class E**:   Throwing the hips to the front 

The goal of this project is to predict the manner in which the participants did 
the exercise, i.e., Class A to E.

More information is available here: [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har) 
(Weight Lifting Exercise Dataset section).

## 2. Executive summary

## 3. Global setup
```{r required packages}
#Loading the required packages
library(knitr); library(caret);         library(rattle);
library(rpart); library(rpart.plot);    library(randomForest);
library(repmis)
```
```{r global knitr setup}
#knitr setup
opts_chunk$set(echo = TRUE) 
```

## 4. Exploratory data analyses

## 5. Prediction Algorithms

### 5.1 Classification trees

### 5.2 Random forests

## 6. Prediction on Testing Set